---
title: "ETC5513_Assignment1"
author: Noor Fatima Moolji
format:
  html:
    theme: cerulean
  pdf: default
execute:
  echo: false
  message: false
  warning: false
---


```{r}
#| label: setup
#| echo: false
#  message: false
# warning: false
library (tinytex)
library (tidyverse)
library (ggplot2)
```

```{r}
# reading data
raw_data <- read.csv("data/36621293_Noor Fatima Moolji.csv")
```

# Introduction
This report analyses data from a personalised marketing campaign, where different levels of marketing expenditure were allocated to customers through different channels. The goal of the analysis is to investigate whether higher marketing spend per customer is related with higher purchase amounts. Understanding the relationship is vital, as it will help evaluate the effectiveness of personalised marketing efforts. These insights will also inform future budget allocations and pricing strategies. 

# Research Question
The report investigates the primary question "Is there a relationship between marketing expenditure and customer purchase amounts". It examines the relationship between marketing spend per customer (independent variable) and purchase amount (dependent variable). This question is explored using data collected from last quarter's marketing campaign.

# Dataset Introduction

```{r}
#| label: data cleaning
#| echo: false
#| include: false
#  message: false
# warning: false

head (raw_data, n=10)
glimpse (raw_data)

# remove duplicates
raw_data <- raw_data |>
    distinct ()

# select relevant variables
raw_data <- raw_data |>
    select(customer_id,marketing_spend,visits, purchase_amount,segment)

# remove rows with missing values
clean_data <- raw_data |>
    drop_na(purchase_amount, marketing_spend)

# convert segment into factor
clean_data <- clean_data |>
    mutate(segment = factor(segment))
```

The dataset records data from a personalised marketing campaign conducted over the last quarter for an online retailer. It captures customer_id,signup_date, marketing_spend, visits, purchase_amount, segment(Consumer, Enterprise or SMB) and notes. The **key variables** are `marketing_spend` and `purchase_amount`, which are the basis for the analysis. 
To prepare the data for this analysis, duplicate records were removed, irrelevant columns were not selected, and rows with missing data were removed. Lastly, the `segment` variable was converted into a factor, so R treats it as a categorical variable.