---
title: "ETC5513_Assignment1"
author: Noor Fatima Moolji
format:
  html:
    theme: cerulean
  pdf: default
execute:
  echo: false
  message: false
  warning: false
---


```{r}
#| label: setup
#| echo: false
#  message: false
# warning: false
library (tinytex)
library (tidyverse)
library (ggplot2)
library(knitr)
library(kableExtra)
library(tidyr)
```

```{r}
# reading data
raw_data <- read.csv("data/36621293_Noor Fatima Moolji.csv")
```

# Introduction
This report analyses data from a personalised marketing campaign in which different levels of marketing expenditure were allocated to customers across various channels. The goal of the analysis is to investigate whether higher marketing spend per customer is associated with higher purchase amounts. Understanding this relationship is vital, as it will help evaluate the effectiveness of personalised marketing efforts. These insights will also inform future budget allocations and pricing strategies.

# Research Question
The report investigates the primary question "Is there a relationship between marketing expenditure and customer purchase amounts". It examines the relationship between marketing spend per customer (independent variable) and purchase amount (dependent variable). This question is explored using data collected from last quarter's marketing campaign.

# Dataset Introduction

```{r}
#| label: data cleaning
#| echo: false
#| include: false
#  message: false
# warning: false

head (raw_data, n=10)
glimpse (raw_data)

# remove duplicates
raw_data <- raw_data |>
    distinct ()

# select relevant variables
raw_data <- raw_data |>
    select(customer_id,marketing_spend, purchase_amount,segment)

# remove rows with missing values
clean_data <- raw_data |>
    drop_na(purchase_amount, marketing_spend)

# convert segment into factor
clean_data <- clean_data |>
    mutate(segment = factor(segment))
```

The dataset records data from a personalised marketing campaign conducted over the last quarter for an online retailer. This data is observational in nature, as it is collected from a campaign. It contains a set of 150 customer records, where each row represents a unique customer. It captures seven variables: customer_id, signup_date, marketing_spend, visits, purchase_amount, segment (Consumer, Enterprise, or SMB), and notes. The key variables are marketing_spend and purchase_amount, which form the basis of the analysis.

To prepare the data for this analysis, the following steps were performed:

1.**Duplicate records were removed** using the distinct() function to ensure each observation is unique, as duplicates could distort the relationship between the two variables.

2. **Irrelevant columns were excluded**,** such as signup_date, visits, and notes, as they do not serve any purpose for the research question or analysis.

3. **Rows with missing data were removed**, as the two key primary variables are important (one is independent, the other is dependent); a row missing either of these values will not be able to contribute anything meaningful to the analysis.

4. **The segment variable was converted into a factor**, as R would otherwise treat it as a string variable. This ensures that R treats Consumer, SMB, and Enterprise as discrete groups, as it will be incorporated into the analysis to gain a deeper understanding of the research question—whether the relationship between marketing spend and purchase amount differs across customer segments, and which segment responds most strongly to marketing expenditure.

## Dataset Description 
The cleaned dataset has `r nrow(clean_data)` rows and `r ncol(clean_data)` variables.

The variables in the dataset are:
1. customer_id : Unique identifier for each customer. data type = string.
2. marketing_spend : The total amount spent on marketing for each customer. data type = double.
3. purchase_amount : The total purchase amount made by each customer. data type= double.
4. segment : The customer segment e.g. SMB, Consumer and Enterprise. data type= factor.

The @tbl-summary-stats below summarises the minimum, median, mean and maximum values for both `marketing_spend` and `purchase_amount`, providing a numerical overview of the two key variables.  
  As we can see, for `marketing_spend`, the mean (155.72) and median (150.86) are close, which indicates that the spend is fairly distributed across customers with no extreme outliers.   
  
  Additionally, for `purchase_amount`, the mean is 123.22 and the median is 124.17, which are again close, indicating a fairly even distribution.

This suggests that our data is not heavily affected by outliers and will be able to reflect genuine patterns.

Notably, the mean of `marketing_spend` is higher than `purchase_amount`, which tells us that, on average, more money is being invested in customers than is returned through purchases.

The @fig-distribution shows how the variables `marketing_spend` and `purchase_amount` are distributed across customers in the campaign. We can see that the `marketing_spend` bars (pink) are fairly evenly distributed across the range, which is consistent with the mean and median seen in the table. Meanwhile, the `purchase_amount` bars (blue) are more concentrated in the lower range (0–200), with a peak around 150 and a sharp drop after 200.

The `purchase_amount` bars are noticeably taller in the lower ranges, while the `marketing_spend` bars are taller in the higher ranges, which visually shows that, on average, more is being spent on marketing than customers are spending on purchases.

```{r}
#| label: tbl-summary-stats
#| tbl-cap: "Summary statistics for marketing spend and purchase amount"
#| echo: false

clean_data |>
  select(marketing_spend, purchase_amount) |>
  summary() |>
  kable ()
```

```{r}
#| label: fig-distribution
#| fig-cap: "Distribution of Marketing Spend and Purchase Amount"
#| echo: false

clean_data  |>
  select(marketing_spend, purchase_amount) |>
  pivot_longer(everything()) |>
  ggplot(aes(x = value, fill = name)) +
  geom_histogram(binwidth = 30, position = "dodge") +
  labs(title = "Distribution of Marketing Spend and Purchase Amount", x = "Amount", y = "")
```

# Results

```{r}
#| label: fig-scatter
#| fig-cap: "Scatter plot of marketing spend VS purchase amount"
#| echo: false

ggplot(clean_data, aes(x=marketing_spend, y=purchase_amount))+
  geom_point(alpha = 0.5, color = "steel blue")+
  geom_smooth(color = "darkred", se = FALSE, method = "lm") +
  labs(title = "Scatter plot of marketing spend VS purchase amount", x = "Marketing Spend", y = "Purchase Amount") +
  theme_minimal()
```


```{r}
#| label: fig-scatter
#| fig-cap: "Scatter plot of marketing spend VS purchase amount, facted by customer segment"
#| echo: false

clean_data|>
  ggplot(aes(x = marketing_spend, y = purchase_amount)) +
  geom_point(alpha = 0.6, color = "steel blue") +
  facet_wrap(~segment) +
  labs(title = "Scatter plot of marketing spend VS purchase amount, facted by customer segment",x = "Marketing Spend", y = "Purchase Amount") +
  theme_minimal()
  theme(legend.position = "none")
```